Perturbing SU($n$)-symmetric interactions

### 2020-01-30 collective_sun/sun_perturbation.tex
improed language in last derivation

Given a system of $N$ spins, we wish to simplify a product of
multi-body operators restricted to the fully symmetric manifold.  For
$p$ multi-body operators, such a product takes the form
\begin{align}
  \P_0 \sp{\prod_{j\in\ZZ_p}
    \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}} \P_0,
  \label{eq:sym_prod_choice}
\end{align}
where $\P_0$ is a projector onto the fully symmetric manifold;
$k\in\C_N\p{M_j}$ is a choice of $M_j$ distinct spins; $O_j\p{k}$
denotes the action of a permutationally-symmetric $M_j$-body operator
acting on spins $k$; and $w_j\p{k}$ is a scalar coefficient for
$O_j\p{k}$.

To simplify the product in \eqref{eq:sym_prod_choice}, we first write
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
  = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} O_j\p{k},
  \label{eq:sym_prod_vec}
\end{align}
where we extend the definitions of $w_j\p{k}$ and $O_j\p{k}$ to all
vectors $k\in\ZZ_N^{M_j}$ by defining symmetric rank-$M_j$ tensors
$w_j$ and $O_j$ (where $O_j$ is an {\it operator-valued} tensor) for
which $w_j\p{k}=O_j\p{k}=0$ if $k$ contains any repeated symbols.

In order to project \eqref{eq:sym_prod_vec} onto the fully symmetric
manifold, we need to collect terms whose operator content is
equivalent up to a permutation of spins.  To this end, we classify
terms in \eqref{eq:sym_prod_vec} by the numbers $g_S$ of indices
shared by all tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.  For
example, a term of the form $w_1\p{a,b,c,d} w_2\p{b,c,d} w_3\p{b}$
with distinct indices $\p{a,b,c,d}\in\C_N\p{4}$ would have
\begin{align}
  g_{\set{1}} = 1,
  &&
  g_{\set{1,2,3}} = 1,
  &&
  g_{\set{1,2}} = 2,
\end{align}
and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
assignment can be represented by the Venn diagram
\begin{align}
  \begin{tikzpicture}]
    \tikzmath{
      \scale = 0.7;
      \rr = 2.5 * \scale;
      \aa = \scale;
      \bb = sqrt(3) * \aa;
      \cc = 2 * \aa;
      \topY = \cc + \rr;
      \crossLeftRightUpper = sqrt(\rr*\rr - \bb*\bb) - \aa;
      \dotTopY = ( \crossLeftRightUpper + \topY ) / 2;
      \dotTopLeftHeight = \crossLeftRightUpper - \scale/5;
      \dotTopLeftXFst = -\rr*0.48 - \scale/3;
      \dotTopLeftXSnd = -\rr*0.48 + \scale/3;
      \labelDist = \scale * 0.8;
      \labelTopY = \topY + \labelDist;
      \labelBottomX = \labelTopY * sqrt(3) / 2;
      \labelBottomY = \labelTopY / 2;
    }
    \draw (0, \cc em) circle [radius = \rr em];
    \draw (-\bb em, -\aa em) circle [radius = \rr em];
    \draw (+\bb em, -\aa em) circle [radius = \rr em];
    \node[draw,fill,shape=circle,scale=0.3] at (0,0) {};
    \node[draw,fill,shape=circle,scale=0.3] at (0, \dotTopY em) {};
    \node[draw,fill,shape=circle,scale=0.3]
    at (\dotTopLeftXFst em, \dotTopLeftHeight em) {};
    \node[draw,fill,shape=circle,scale=0.3]
    at (\dotTopLeftXSnd em, \dotTopLeftHeight em) {};
    \node at (0, \labelTopY em) {1};
    \node at (-\labelBottomX em, -\labelBottomY em) {2};
    \node at (+\labelBottomX em, -\labelBottomY em) {3};
  \end{tikzpicture},
  \label{eq:venn_diagram}
\end{align}
where $g_S$ is determined the number of dots at the intersection of
the circles $j\in S$.  We denote an assignment of distinct indices
according to choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ by $g$, where
$\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets) of $\ZZ_p$.
Given a vector $\m M$ of the tensor ranks $M_j$, we then denote the
set of all valid index assignments $g$ by $\G\p{\m M}$.  For
consistency, all valid index assignments $g\in\G\p{\m M}$ must assign
exactly $M_j$ indices to the rank-$M_j$ tensor $w_j$, i.e.~
\begin{align}
  \sum_{S\in\PP\p{\ZZ_p}\,:\,j\in S} g_S
  = \sum_{R\in\PP\p{\ZZ_p\setminus\set{j}}} g_{\set{j}\cup R}
  = M_j
\end{align}
for all $j\in\ZZ_p$.  In order to make the index assignments
determined by choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ unique, we
also enforce that $g_{\set{}}=0$ for all $g\in\G\p{\m M}$, which is
equivalent to saying that every dot in the corresponding Venn diagram
must lie within some circle.  The set of valid index assignments
$\G\p{\m M}$ is then essentially the set of all Venn diagrams of the
form in \eqref{eq:venn_diagram} with $p=\rank\p{\m M}$ circles and a
total of $M_j$ dots within circle $j$.

A classification of terms in \eqref{eq:sym_prod_vec} by index
assignments $g\in\G\p{\m M}$ allows us to expand
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} O_j\p{k}
  \EQFS \sum_{g\in\G\p{\m M}} \S\p{g} w\p{g} O\p{g},
  \label{eq:sym_prod_group_start}
\end{align}
where $O\p{g}$ is an operator acquired assigning indices to the
tensors $O_j$ in a manner consistent with $g$; $w\p{g}$ is a scalar
acquired by summing over all indices assigned to the tensors $w_j$
according to $g$; and $\S\p{g}$ is a symmetry factor accounting for
the number of equivalent ways to assign indices according to $g$.

In order to write out the factors in \eqref{eq:sym_prod_group_start}
explicitly, we define the set of values that the distinct indices
assigned according to $g$ can take,
\begin{align}
  \C_N\p{g}
  \equiv \set{ k \in \bigotimes_{S\in\PP\p{\ZZ_p}} \C_N\p{g_S}
    : \t{all elements of $k$ are distinct} }
  \simeq \C_N\p{\sum_{S\in\PP\p{\ZZ_p}}g_S},
\end{align}
where each $k\in\C_N\p{g}$ has the natural decomposition
$k=\p{k_S:S\in\PP\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$, and we define
the restriction of $k\in\C_N\p{g}$ to values associated with the
tensors $w_j,O_j$,
\begin{align}
  k_j \equiv \bigcup_{S\in\PP\p{\ZZ_p} : j\in S} k_S
  = \bigcup_{R\in\PP\p{\ZZ_p\setminus\set{j}}} k_{\set{j}\cup R},
\end{align}
where the union of lists $k_S$ denotes a concatenation of those lists.  We then have that
\begin{align}
  w\p{g} \equiv \sum_{k\in\C_N\p{g}} \prod_{j\in\ZZ_p} w_j\p{k_j},
  &&
  O\p{g}
  \equiv \prod_{j\in\ZZ_p} O_j\p{k_j}~\t{for some}~k\in\C_N\p{g},
\end{align}
and
\begin{align}
  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} { M_j \choose g_S : j\in S }}
  \sp{\prod_{S\in\PP\p{\ZZ_p}} \p{g_S!}^\abs{S}}
  = \prod_{j\in\ZZ_p} M_j!,
  \label{eq:sym_prod_symmetry}
\end{align}
where the first product in \eqref{eq:sym_prod_symmetry} contains multinomial coefficients
\begin{align}
  { M \choose a_1, a_2, \cdots, a_m }
  \equiv \f{M!}{a_1!a_2!\cdots a_m!},
\end{align}
that account for the number of ways to partition the indices of each
$w_j,O_j$ into sets of shared indices according to $g_S$ for all
$j\in S$, and the second product in \eqref{eq:sym_prod_symmetry}
accounts for the number of ways to permute each set of shared indices
on every tensor.  The choice of $k\in\C_N\p{g}$ in the definition of
$O\p{g}$ is irrelevant after projection onto the fully symmetric
manifold.  Altogether, we have that
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
  \EQFS \sum_{g\in\G\p{\m M}} w\p{g} O\p{g}.
  \label{eq:sym_prod_group}
\end{align}

>>>>

Given a system of $N$ spins, we wish to project a product of
multi-body operators onto the fully symmetric manifold.  For $p$
multi-body operators, the projection of such a product takes the form
\begin{align}
  \P_0 \sp{\prod_{j\in\ZZ_p}
    \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}} \P_0,
  \label{eq:sym_prod_proj}
\end{align}
where $\P_0$ is a projector onto the fully symmetric manifold;
$k\in\C_N\p{M_j}$ is a choice of $M_j$ distinct spins; $O_j\p{k}$
denotes the action of an $M_j$-body operator on spins $k$ that is
invariant under permutations of $k$; and $w_j\p{k}$ is a scalar
coefficient for $O_j\p{k}$.

To simplify the product in \eqref{eq:sym_prod_proj}, we first write
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
  = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}}
  w_j\p{k} O_j\p{k},
  \label{eq:sym_prod_vec}
\end{align}
where we extend the definitions of $w_j\p{k}$ and $O_j\p{k}$ to all
vectors $k\in\ZZ_N^{M_j}$ by defining symmetric rank-$M_j$ tensors
$w_j$ and $O_j$ (where $O_j$ is an {\it operator-valued} tensor) for
which $w_j\p{k}=O_j\p{k}=0$ if $k$ contains any repeated symbols.  We
then collect terms whose operator content is equivalent up to a
permutation of all spins.  To this end, we classify terms in
\eqref{eq:sym_prod_vec} by the numbers $g_S$ of indices shared by all
tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.  For example, a term of
the form $w_1\p{a,b,c,d} w_2\p{b,c,d} w_3\p{b}$ with distinct indices
$\p{a,b,c,d}\in\C_N\p{4}$ would have
\begin{align}
  g_{\set{1}} = 1,
  &&
  g_{\set{1,2,3}} = 1,
  &&
  g_{\set{1,2}} = 2,
\end{align}
and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
assignment can be represented by the Venn diagram
\begin{align}
  \begin{tikzpicture}]
    \tikzmath{
      \scale = 0.7;
      \rr = 2.5 * \scale;
      \aa = \scale;
      \bb = sqrt(3) * \aa;
      \cc = 2 * \aa;
      \topY = \cc + \rr;
      \crossLeftRightUpper = sqrt(\rr*\rr - \bb*\bb) - \aa;
      \dotTopY = ( \crossLeftRightUpper + \topY ) / 2;
      \dotTopLeftHeight = \crossLeftRightUpper - \scale/5;
      \dotTopLeftXFst = -\rr*0.48 - \scale/3;
      \dotTopLeftXSnd = -\rr*0.48 + \scale/3;
      \labelDist = \scale * 0.8;
      \labelTopY = \topY + \labelDist;
      \labelBottomX = \labelTopY * sqrt(3) / 2;
      \labelBottomY = \labelTopY / 2;
    }
    \draw (0, \cc em) circle [radius = \rr em];
    \draw (-\bb em, -\aa em) circle [radius = \rr em];
    \draw (+\bb em, -\aa em) circle [radius = \rr em];
    \node[draw,fill,shape=circle,scale=0.3] at (0,0) {};
    \node[draw,fill,shape=circle,scale=0.3] at (0, \dotTopY em) {};
    \node[draw,fill,shape=circle,scale=0.3]
    at (\dotTopLeftXFst em, \dotTopLeftHeight em) {};
    \node[draw,fill,shape=circle,scale=0.3]
    at (\dotTopLeftXSnd em, \dotTopLeftHeight em) {};
    \node at (0, \labelTopY em) {1};
    \node at (-\labelBottomX em, -\labelBottomY em) {2};
    \node at (+\labelBottomX em, -\labelBottomY em) {3};
  \end{tikzpicture},
  \label{eq:venn_diagram}
\end{align}
where $g_S$ is determined the number of dots at the intersection of
the circles $j\in S$.  We denote an assignment of distinct indices
according to choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ by $g$, where
$\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets) of $\ZZ_p$.
Given a vector $\m M$ of the tensor ranks $M_j$, we then denote the
set of all valid index assignments $g$ by $\G\p{\m M}$.  For
consistency, all valid index assignments $g\in\G\p{\m M}$ must assign
exactly $M_j$ indices to the rank-$M_j$ tensor $w_j$, i.e.~
\begin{align}
  \sum_{S\in\PP\p{\ZZ_p}\,:\,j\in S} g_S
  = \sum_{R\in\PP\p{\ZZ_p\setminus\set{j}}} g_{\set{j}\cup R}
  = M_j
\end{align}
for all $j\in\ZZ_p$.  In order to make the index assignments
determined by choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ unique, we
also enforce that $g_{\set{}}=0$ for all $g\in\G\p{\m M}$, which is
equivalent to saying that every dot in the corresponding Venn diagram
must lie within some circle.  The set of valid index assignments
$\G\p{\m M}$ is then essentially the set of all Venn diagrams of the
form in \eqref{eq:venn_diagram} with $p=\rank\p{\m M}$ circles and a
total of $M_j$ dots within circle $j$.

A classification of terms in \eqref{eq:sym_prod_vec} by index
assignments $g\in\G\p{\m M}$ allows us to expand
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} O_j\p{k}
  \EQFS \sum_{g\in\G\p{\m M}} \S\p{g} w\p{g} O\p{g},
  \label{eq:sym_prod_group_start}
\end{align}
where $O\p{g}$ is an operator acquired by assigning indices to the
tensors $O_j$ in a manner consistent with $g$; $w\p{g}$ is a scalar
acquired by summing over all indices assigned to the tensors $w_j$
according to $g$; and $\S\p{g}$ is a symmetry factor accounting for
the number of equivalent ways to assign indices according to $g$.

In order to write out the factors in \eqref{eq:sym_prod_group_start}
explicitly, we define
\begin{align}
  \abs{g} \equiv \sum_{S\in\PP\p{\ZZ_p}} g_S,
\end{align}
which is equivalent to the total number of dots in the Venn diagram
associated with $g$, and identify the set of values that the distinct
indices assigned according to $g$ can take:
\begin{align}
  \C_N\p{g}
  \equiv \set{ k \in \bigotimes_{S\in\PP\p{\ZZ_p}} \C_N\p{g_S}
    : \t{all elements of $k$ are distinct} }
  \simeq \C_N\p{\abs{g}},
\end{align}
where each $k\in\C_N\p{g}$ has the natural decomposition
$k=\p{k_S:S\in\PP\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$.  We then define
the restriction of $k\in\C_N\p{g}$ to values associated with $w_j$ and
$O_j$:
\begin{align}
  k_j \equiv \bigcup_{S\in\PP\p{\ZZ_p} : j\in S} k_S
  = \bigcup_{R\in\PP\p{\ZZ_p\setminus\set{j}}} k_{\set{j}\cup R},
\end{align}
where the union of lists $k_S$ denotes a concatenation of those lists.
We then have that
\begin{align}
  w\p{g} \equiv \sum_{k\in\C_N\p{g}} \prod_{j\in\ZZ_p} w_j\p{k_j},
  &&
  O\p{g} \equiv \prod_{j\in\ZZ_p}
  O_j\p{\ell_j}~\t{for some}~\ell\in\C_N\p{g},
\end{align}
where the choice of $\ell\in\C_N\p{g}$ does not matter after
projection onto fully symmetric manifold, and
\begin{align}
  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} { M_j \choose g_S : j\in S }}
  \sp{\prod_{S\in\PP\p{\ZZ_p}} \p{g_S!}^\abs{S}}
  = \prod_{j\in\ZZ_p} M_j!,
  \label{eq:sym_prod_symmetry}
\end{align}
where the first product in \eqref{eq:sym_prod_symmetry} contains
multinomial coefficients
\begin{align}
  { M \choose a_1, a_2, \cdots, a_m }
  \equiv \f{M!}{a_1!a_2!\cdots a_m!},
\end{align}
that account for the number of ways to partition the indices of each
$w_j,O_j$ into sets of shared indices according to $g_S$ for all
$j\in S$, and the second product in \eqref{eq:sym_prod_symmetry}
accounts for the number of ways to permute each set of shared indices
on every tensor.  Altogether, the symmetry factors $\S\p{g}$ turn out
to be independent of $g$, and cancel out exactly with the $\sim M_j!$
prefactors in \eqref{eq:sym_prod_vec}, so
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
  \EQFS \sum_{g\in\G\p{\m M}} w\p{g} O\p{g}.
  \label{eq:sym_prod_group}
\end{align}
